---
title: "p8105 Homework 6"
author: "Elaine Yanxi Chen"
date: "`r Sys.Date()`"
output: github_document
---

## Packages and settings

First we load the packages necessary to knit this document.

```{r packages and settings, message = FALSE}
library(tidyverse)
library(mgcv)
library(modelr)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```


## Problem 2: Homicides in Baltimore, MD

### Data Import

After loading the dataset, we want to do some initial cleaning and tidying following these criteria below:

* We want to load the data on homicides in 50 large U.S. cities. 
* We want to create a `city_state` variable and a binary variable indicating whether the homicide is solved.
* Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO since they do not report victim race. Omit Tulsa, AL because this is a data entry mistake.
* We will limit the analysis for whom `victim_race` is `white` or `black`.
* Make sure that `victim_age` is a numeric variable.

```{r homicide_df}
homicide_df = 
  read_csv("data/homicide-data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    city_state = str_c(city, state, sep = ","),
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")
  ) %>% 
  filter(
    !city_state %in% c("Dallas,TX", "Phoenix,AZ", "Kansas City,MO", "Tulsa,AL"),
    victim_race %in% c("White", "Black"))
```


### Baltimore, MD

We are interested in using the `glm` function to fit a logistic regression model with resolved vs unresolved as the outcome, and victim age, sex, and race as predictors.

To do so, we will first generate a dataset containing these variables only.

```{r baltimore_df}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore,MD") %>% 
  select(resolved, victim_age, victim_race, victim_sex)
```

Now we will run the logistic regression model using `glm` and save the output as an R object.

```{r fit_logistic}
fit_logistic = 
  baltimore_df %>% 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) 
```

```{r clean glm_baltimore}
glm_baltimore = 
  fit_logistic %>% 
  broom::tidy() %>% 
  janitor::clean_names() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std_error),
    CI_upper = exp(estimate + 1.96 * std_error)) %>% 
  select(-c(std_error, statistic)) 

glm_baltimore %>% 
  knitr::kable(digits = 3)
```

After running the logistic regression model, we want to obtain the estimate and the confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed. 

```{r victim sex}
glm_baltimore %>% 
  filter(term == "victim_sexMale") %>% 
  select(OR:CI_upper) %>% 
  knitr::kable(
    digits = 3,
    col.names = c("Odds Ratio", "Lower 95% CI", "Upper 95% CI"))
```

We see that keeping all other variables fixed, homicides in which the victim is male are 57.4% less likely to be resolved than those in which the victim is female. We are 95% confident that the true odds ratio for the association between victim's sex and resolved homicides lies between 0.325 and 0.558. 


### Iterating `glm` for each city

Now we want to run `glm` for each of the cities in the datset.

First, we will create a dataset containing all cities but only with variables of interest.

```{r}
homicide_glm_df =
  homicide_df %>% 
  select(city_state, resolved, victim_age, victim_race, victim_sex)
```

Now we will map the `glm` function across all cities, tidy the output, and finally filter for `victim_sex` and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. 

```{r}
nest_glm_city =
  homicide_glm_df %>% 
  nest(df = -city_state) %>% 
  mutate(
    models = map(.x = df, ~glm(resolved ~ victim_age + victim_sex + victim_race, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results)

glm_city =
  nest_glm_city %>% 
  janitor::clean_names() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std_error),
    CI_upper = exp(estimate + 1.96 * std_error)) %>% 
  select(-c(std_error, statistic)) %>% 
  filter(term == "victim_sexMale") %>% 
  select(city_state, OR:CI_upper) 

glm_city %>% 
  knitr::kable(
    digits = 3,
    col.names = c("City, State", "Odds Ratio", "Lower 95% CI", "Upper 95% CI"))
```


### Plot for estimated ORs and CIs for each city

Lastly, we want to create a plot showing the estimated ORs and CIs for each city. We will also show it in a way so that cities are organized according to the their estimated OR. 

```{r}
glm_city %>% 
  mutate(
    city_state = fct_reorder(city_state, OR, .desc = TRUE)
  ) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point(alpha = 0.5) + 
  theme(axis.text.x = element_text(angle = 80, hjust = 1)) +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper),
                alpha = 0.5) +
  labs(
    title = "Estimated Odds Ratios and 95% Confidence Intervals of Solved Homicides and Victim Sex for Large U.S.Cities",
    x = "City, State",
    y = "Estimates Odds Ratios (95% Confidence Intervals)"
  )
```


## Problem 3: Child's Birthweight

In this problem, we will analyze data gathered on child's birthweight and several variables.

We will first load the data and inspect it, before cleanning it for regression analysis.

```{r}
childbw_df = 
  read_csv("data/birthweight.csv") %>% 
  janitor::clean_names()
```

